### 2021-11-09 pres/group-meeting.Rmd
save

---
title: "Group meeting"
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, width = 70)
knitr::opts_chunk$set(fig.align = 'center', dev = "svg", out.width = "70%",
                      echo = FALSE, comment = "", fig.width = 5, global.par = TRUE)
```

class: title-slide center middle inverse

<br>

# Global ancestry proportions and frequencies from the UK Biobank

<br>

## Florian Privé

---

## Goal

<br>

#### Infer ancestry proportions using allele frequencies only

---

## Main challenges

<br>

- Need some (precise) frequencies from many global populations

--

- Need a method to infer ancestry proportions.    
    We can minimize the following problem, subject to $w \ge 0$ and $\sum_k w_k = 1$ (convex combination):

    $$\min_w \sum_{j=1}^M \left( f_j^{(0)} - \sum_{k=1}^K f_j^{(k)} w_k \right)^2 ~,$$
    where $M$ is the number of variants, $K$ the number of populations, $f_j^{(k)}$ is the frequency of variant $j$ in population $k$, and $f_j^{(0)}$ is the frequency of variant $j$ in the global population of interest.

---

## A set of global frequencies

<br>

- The 1000 Genomes data is too small, and missing some parts of the world

- HGDP, EGDP or SGDP are too small as well

--

<br>

- therefore I used the UK Biobank, which includes individuals from a bit of everywhere + secret recipe for ancestry estimation -> frequencies for 20 global populations

    <br>

    ```{r, out.width = "100%"}
    knitr::include_graphics("group-sizes.jpg")
    ```

---

## Verification

<br>

Computing $F_{ST}$ with closest population from the 1000G data:

<br>

```{r, out.width = "100%"}
knitr::include_graphics("group-fst.jpg")
```

---

## Controls (1/2)

<br>

- Biobank Japan:
    ```{r, out.width = "50%"}
    knitr::include_graphics("res-bbj.jpg")
    ```

<br>

- FinnGen:
    ```{r, out.width = "57%"}
    knitr::include_graphics("res-finngen.jpg")
    ```
---

## Controls (2/2)

<br>

- UK Biobank:
    ```{r, out.width = "60%"}
    knitr::include_graphics("res-ukbb.jpg")
    ```

---

## Other results (1/3)

- The COVID-19 Host Genetics Initiative ("European" subset):
    ```{r, out.width = "42%"}
    knitr::include_graphics("res-covid-eur.jpg")
    ```
- All (except 23andMe):
    ```{r, out.width = "48%"}
    knitr::include_graphics("res-covid-no23andme.jpg")
    ```

---

## Other results (2/3)

- Epilepsy (EUR + EAS + AFR):
    ```{r, out.width = "52%"}
    knitr::include_graphics("res-epilepsy.jpg")
    ```

<br>

> Furthermore, 531 cases of Asian descent, and 147 cases of African descent were included through a meta-analysis.

---

## Other results (3/3)

- PAGE consortium:
    ```{r, out.width = "45%"}
    knitr::include_graphics("res-page.jpg")
    ```

And the "truth" according to their paper:

```{r, out.width = "100%"}
knitr::include_graphics("page-truth.jpg")
```

---

## Conclusions

- Works well, but still not perfect (e.g. for single populations)

- Missing populations? (e.g. Native Americans, Native Hawaiian)

- Could add Ashkenazi Jewish group, but still a bit hesitant doing so

--

<br>

## Ideas for improvement

- Increase sample size of groups for more precise allele frequencies

- Use millions of variants (for now, only 500K genotyped variants of UKBB)

- Filter bad variants in the UKBB

- Give more weight to more ancestry-informative variants? to rarer variants?